RDF Validation of Metadata \\ on Person-Level and Aggregated Data

### 2015-03-11 paper/paper7/paper7-appendix.tex
evaluation

	\item \textbf{{\em SKOS-C-LANGUAGE-TAG-CARDINALITY-01\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Omitted or Invalid Language Tags}:}}
Some controlled vocabularies contain literals in natural language, but without information what language has actually been used. Language tags might also not conform to language standards, such as RFC 3066. 
Iteration over all triples in the vocabulary that have a predicate which is a (subclass of) \emph{rdfs:label} or \emph{skos:note}. 
Severity level: INFO

  \item \textbf{{\em SKOS-C-LANGUAGE-TAG-CARDINALITY-02\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Incomplete Language Coverage}:}}
	Some concepts in a thesaurus are labeled in only one language, some in multiple languages. It may be desirable to have each concept labeled in each of the languages that also are used on the other concepts. This is not always possible, but incompleteness of language coverage for some concepts can indicate shortcomings of the vocabulary. 
	Severity level: INFO

	\item \textbf{{\em SKOS-C-LANGUAGE-TAG-CARDINALITY-03\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - No Common Language}:}}
	Checks if all concepts have at least one common language, i.e. they have assigned at least one literal in the same language. 
	Severity level: INFO

@@

		\item Implementation: Iteration over all concepts in the vocabulary and returning that don't have associated resources using (subproperties of) \emph{skos:semanticRelation}. 

@@

@@

	  \item Severity level: SHOULD

@@

	As stated in the SKOS reference document, \emph{skos:broader} and \emph{skos:narrower} are not transitive properties. However, they are subproperties of \emph{skos:broaderTransitive} and \emph{skos:narrowerTransitive} which enables inference of a "transitive closure". This, in fact, leaves it up to the user to interpret wheter a vocabulary's hierarchical structure is seen as transitive or not. In the former case, this check can be useful. It finds pairs of concepts (A,B) that are directly hierarchically related but there also exits an hierarchical path through a concept C that connects A and B. 
	Severity level: INFO

@@

\begin{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-01\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Undocumented Concepts}:}}
	Undocumented Concepts:
  The SKOS standard defines a number of properties useful for documenting the meaning of the concepts in a thesaurus also in a human-readable form. Intense use of these properties leads to a well-documented thesaurus which should also improve its quality.  
	Iteration over all concepts in the vocabulary and find those not using one of \emph{skos:note}, \emph{skos:changeNote}, \emph{skos:definition}, \emph{skos:editorialNote}, \emph{skos:example}, \emph{skos:historyNote}, or \emph{skos:scopeNote}.
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-02\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Overlapping Labels}:}}
	Overlapping Labels:
	This is a generalization of a recommendation in the SKOS primer, that “no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme”. This could indicate missing disambiguation information and thus lead to problems in autocompletion application. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-03\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Missing Labels}:}}
	Missing Labels:
	To make the vocabulary more convenient for humans to use, instances of SKOS classes (Concept, ConceptScheme, Collection) should be labeled using e.g., \emph{skos:prefLabel}, \emph{altLabel}, \emph{rdfs:label}, \emph{dc:title}. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-04\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Unprintable Characters in Labels}:}}
	Unprintable Characters in Labels:
	\emph{pref/alt/hiddenlabels} contain characters that are not alphanumeric characters or blanks.
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-05\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Empty Labels}:}}
	Empty Labels:
	Labels also need to contain textual information to be useful, thus we find all SKOS labels with length 0 (after removing whitespaces). 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-06\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Ambiguous Notation References}:}}
	Ambiguous Notation References:
	Concepts within the same concept scheme should not have identlical \emph{skos:notation} literals. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
\end{itemize}

@@

>>>>

	\item \textbf{{\em SKOS-C-LANGUAGE-TAG-CARDINALITY-01\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Omitted or Invalid Language Tags}:}}
	Omitted or Invalid Language Tags:
Some controlled vocabularies contain literals in natural language, but without information what language has actually been used. Language tags might also not conform to language standards, such as RFC 3066. 
\begin{itemize}
	\item Implementation: Iteration over all triples in the vocabulary that have a predicate which is a (subclass of) \emph{rdfs:label} or \emph{skos:note}. 
  \item Severity level: WARNING
\end{itemize}

  \item \textbf{{\em SKOS-C-LANGUAGE-TAG-CARDINALITY-02\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Incomplete Language Coverage}:}}
	Incomplete Language Coverage:
	Some concepts in a thesaurus are labeled in only one language, some in multiple languages. It may be desirable to have each concept labeled in each of the languages that also are used on the other concepts. This is not always possible, but incompleteness of language coverage for some concepts can indicate shortcomings of the vocabulary. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}

	\item \textbf{{\em SKOS-C-LANGUAGE-TAG-CARDINALITY-03\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - No Common Language}:}}
	No Common Language:
	Checks if all concepts have at least one common language, i.e. they have assigned at least one literal in the same language. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}

@@

		\item Implementation: Iteration over all concepts in the vocabulary and returning that don't have associated resources using (sub-properties of) \emph{skos:semanticRelation}. 
		\item Severity level: WARNING

@@

		\item Severity level: WARNING

@@

	  \item Severity level: WARNING

@@

	As stated in the SKOS reference document, \emph{skos:broader} and \emph{skos:narrower} are not transitive properties. However, they are sub-properties of \emph{skos:broaderTransitive} and \emph{skos:narrowerTransitive} which enables inference of a "transitive closure". This, in fact, leaves it up to the user to interpret whether a vocabulary's hierarchical structure is seen as transitive or not. In the former case, this check can be useful. It finds pairs of concepts (A,B) that are directly hierarchically related but there also exits an hierarchical path through a concept C that connects A and B. 

@@

\begin{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-01\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Undocumented Concepts}:}}
	Undocumented Concepts:
  The SKOS standard defines a number of properties useful for documenting the meaning of the concepts in a thesaurus also in a human-readable form. Intense use of these properties leads to a well-documented thesaurus which should also improve its quality.  
	\begin{itemize}
	  \item Implementation:
		Iteration over all concepts in the vocabulary and find those not using one of \emph{skos:note}, \emph{skos:changeNote}, \emph{skos:definition}, \emph{skos:editorialNote}, \emph{skos:example}, \emph{skos:historyNote}, or \emph{skos:scopeNote}.
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-02\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Overlapping Labels}:}}
	Overlapping Labels:
	This is a generalization of a recommendation in the SKOS primer, that “no two concepts have the same preferred lexical label in a given language when they belong to the same concept scheme”. This could indicate missing disambiguation information and thus lead to problems in autocompletion application. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-03\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Missing Labels}:}}
	Missing Labels:
	To make the vocabulary more convenient for humans to use, instances of SKOS classes (Concept, ConceptScheme, Collection) should be labeled using e.g., \emph{skos:prefLabel}, \emph{altLabel}, \emph{rdfs:label}, \emph{dc:title}. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-04\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Unprintable Characters in Labels}:}}
	Unprintable Characters in Labels:
	\emph{pref/alt/hiddenlabels} contain characters that are not alphanumeric characters or blanks.
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-05\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Empty Labels}:}}
	Empty Labels:
	Labels also need to contain textual information to be useful, thus we find all SKOS labels with length 0 (after removing whitespaces). 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
	\item \textbf{{\em SKOS-C-LABELING-AND-DOCUMENTATION-06\footnote{Corresponds to qSKOS Quality Issues - Labeling and Documentation Issues - Ambiguous Notation References}:}}
	Ambiguous Notation References:
	Concepts within the same concept scheme should not have identical \emph{skos:notation} literals. 
	\begin{itemize}
		\item Severity level: INFO
	\end{itemize}
\end{itemize}

@@

\subsection{\underline{Linked Data}}

\begin{itemize}
	\item \textbf{\em R-HTTP-URI-SCHEME-VIOLATION\footnote{Corresponds to qSKOS Quality Issues - Linked Data Specific Issues - HTTP URI Scheme Violation}}: 
	In the context of Linked Data, we restrict ourselves to using HTTP URIs only and avoid other URI schemes such as URNs and DOIs.
	\begin{itemize}
		\item Severity level: ERROR
	\end{itemize}
\end{itemize}

### 2015-03-11 paper/paper7/paper7.tex
evaluation

\textbf{Unique Identification.}
It is often useful to declare a given (data) property as the \emph{primary key} (\emph{R-226}) of a class, so that a system can enforce uniqueness and also automatically build URIs from user inputs and imported data. 
In \emph{Disco}, resources are uniquely identified by the property \emph{adms:identifier},
which is therefore inverse-functional
$(\ms{funct identifier}\sp{\overline{\ }})$,
i.e. for each \emph{rdfs:Resource x}, there can be at most one distinct \emph{rdfs:Resource y} such that \emph{y} is connected by \emph{adms:identifier$\sp{\overline{\ }}$} to \emph{x}.
Keys, however, are even more general than inverse-functional properties (\emph{R-58}),
as a key can be a data, an object property, or a chain of properties \cite{Schneider2009}.
For this generalization purposes, as there are different sorts of key, and as keys can lead to undecidability, 
DL is extended with \emph{key boxes} and a special \emph{keyfor} construct (\ms{identifier \ms{keyfor} Resource}) \cite{Lutz2005}.

@@

In \emph{Missy}, we provide comprehensive Linked Data services\footnote{\textcolor{red}{\url{http://www.gesis.org/missy/eu/general-information/linked-data/}}} like a \textcolor{red}{triple store} and diverse RDF exports of person-level metadata conforming to the \emph{Disco} vocabulary in multiple concrete syntaxes (e.g. turtle).

SPARQL is generally seen as the method of choice to validate RDF data according to certain constraints.
We use \emph{SPIN}\footnote{\url{http://spinrdf.org/}}, 
a SPARQL-based way to formulate and check constraints, as basis to define a
validation environment (available at \url{http://purl.org/net/rdfval-demo}) to validate RDF data according to constraints expressed my arbitrary constraint languages like Shape Expressions\footnote{\url{http://www.w3.org/Submission/shex-primer/}} and the Web Ontology Language \cite{BoschEckert2014-2}.
The \emph{RDF Validator} also validates RDF data conforming to SPARQL constraints to ensure correct syntax and semantics of vocabularies such as \emph{Disco, Data Cube, PHDD, DCAT, SKOS, and XKOS}.
The SPIN engine checks for each resource if it satisfies all constraints (associated with its assigned classes) and generates a result RDF graph containing information about all constraint violations.
We define one SPIN construct template for each constraint type (corresponding to a requirement) and vocabulary-specific constraint\footnote{For details about the validation environment see \cite{BoschEckert2014-2}}.
A SPIN construct template contains a SPARQL CONSTRUCT query which generates constraint violation triples indicating the subject, the properties causing constraint violations, and the reason why constraint violations have been raised.
A SPIN construct template creates constraint violation triples if all triple patterns within the SPARQL WHERE clause match.

@@

We evaluated multiple aggregated and person-level data sets

the most important constraints and constraint types

- Data Cube: XXXXX
- 3 thesauri: The Thesaurus for the Social Sciences (TheSoz), STW Thesaurus for Economics of the ZBW, and the AGROVOC thesaurus of FAO
- Disco: 5 series like EU-SILC 
- PHDD
- XKOS

in order to ensure high quality of existing large real world data sets

Exported from Missy, we validated \emph{Disco} person-level data containing overall XX resources (5 series, XX studies, XX data sets, XX theoretical concepts, XX variables, XX questions, XX codes, XX category statistics (frequencies), and XX summary statistics).

triple store\footnote{\textcolor{red}{\url{http://multiweb.gesis.org/openrdf-workbench/}}}

The correctness of $\mathcal{C}_{C}$ constraints, i.e., the gold standard has been designed by a SBE domain expert.

>>>>

\textbf{Unique Identification.}
It is often useful to declare a given (data) property as the \emph{primary key} (\emph{R-226}) of a class, so that a system can enforce uniqueness and also automatically build URIs from user inputs and imported data. 
In \emph{Disco}, resources are uniquely identified by the property \emph{adms:identifier},
which is therefore inverse-functional
$(\ms{funct identifier}\sp{\overline{\ }})$,
i.e. for each \emph{rdfs:Resource x}, there can be at most one distinct \emph{rdfs:Resource y} such that \emph{y} is connected by \emph{adms:identifier$\sp{\overline{\ }}$} to \emph{x}.
Keys, however, are even more general than inverse-functional properties (\emph{R-58}),
as a key can be a data, an object property, or a chain of properties \cite{Schneider2009}.
For this generalization purposes, as there are different sorts of key, and as keys can lead to undecidability, 
DL is extended with \emph{key boxes} and a special \emph{keyfor} construct (\ms{identifier \ms{keyfor} Resource}) \cite{Lutz2005}.
OWL 2 \emph{hasKey} implements \emph{keyfor} and can therefore be used to identify resources uniquely, to merge resources with identical key property values, and to recognize constraint violations.
%OWL 2 hasKey can be used to identify resources uniquely
%
%We used Protégé 5.
%
%example: owl:Thing owl:hasKey ( :hasSSN ) . :Peter :hasSSN "123-45-6789" .
%:Peter_Griffin :hasSSN "123-45-6789" .
%
%We use the predefined Reasoner HermiT.
%
%:Peter and :Peter_Griffin are derived as identical from the reasoner as they have the same value for :hasSSN.
%
%hasKey can be used to merge resources and to recognize constraint violations.
%
%Alternative to OWL 2 hasKey: concise bounded description (proposal of Dan)
%
%action: explore CBD document solution might be using both approaches
%
%benefit: additional identification to URI / IRI benefit: compound properties can be used as key

@@

In \emph{Missy}, we provide comprehensive Linked Data services\footnote{\textcolor{red}{\url{http://www.gesis.org/missy/eu/general-information/linked-data/}}} like a triple store and diverse RDF exports of person-level metadata conforming to the \emph{Disco} vocabulary in multiple concrete syntaxes (e.g. turtle).

SPARQL is generally seen as the method of choice to validate RDF data according to certain constraints.
We use \emph{SPIN}\footnote{\url{http://spinrdf.org/}}, 
a SPARQL-based way to formulate and check constraints, as basis to define a
validation environment (available at \url{http://purl.org/net/rdfval-demo}) to validate RDF data according to constraints expressed my arbitrary constraint languages like Shape Expressions\footnote{\url{http://www.w3.org/Submission/shex-primer/}} and the Web Ontology Language \cite{BoschEckert2014-2}.
The \emph{RDF Validator} also validates RDF data conforming to SPARQL constraints to ensure correct syntax and semantics of vocabularies such as \emph{Disco, Data Cube, PHDD, DCAT, SKOS, and XKOS}\footnote{\textcolor{red}{SPIN mappings: \url{https://github.com/boschthomas/rdf-validation/SPIN}}}.
The SPIN engine checks for each resource if it satisfies all constraints (associated with its assigned classes) and generates a result RDF graph containing information about all constraint violations.
We define one SPIN construct template for each constraint type (corresponding to a requirement) and vocabulary-specific constraint\footnote{For details about the \emph{RDF Validator} see \cite{BoschEckert2014-2}}.
A SPIN construct template contains a SPARQL CONSTRUCT query which generates constraint violation triples indicating the subject, the properties causing constraint violations, and the reason why constraint violations have been raised.
A SPIN construct template creates constraint violation triples if all triple patterns within the SPARQL WHERE clause match.

@@

We used $\mathcal{C}_{C}$ and $\mathcal{C}_{T}$ constraints 
to evaluate the metadata and data quality of existing large real world aggregated and person-level data sets.
We provide a triple store\footnote{\textcolor{red}{\url{http://multiweb.gesis.org/openrdf-workbench/}}}, containing these data sets, 
and the constraints\footnote{\url{https://github.com/boschthomas/rdf-validation/constraints}} in form of SPARQL CONSTRUCT queries.
We validated 
(1) \emph{QB} data sets (\textcolor{red}{XX}),
(2) \emph{Disco} data sets (overall \textcolor{red}{XX} resources exported from \emph{Missy} - 5 series, \textcolor{red}{XX} studies, \textcolor{red}{XX} data sets, \textcolor{red}{XX} theoretical concepts, \textcolor{red}{XX} variables, \textcolor{red}{XX} questions, \textcolor{red}{XX} codes, \textcolor{red}{XX} category statistics (frequencies), and \textcolor{red}{XX} summary statistics),
(3) \emph{DCAT} data sets (\textcolor{red}{XX}),
(4) \emph{PHDD} data sets (\textcolor{red}{XX}),
(5) \emph{SKOS} data sets (The Thesaurus for the Social Sciences (TheSoz), the STW Thesaurus for Economics of the ZBW, and the AGROVOC thesaurus of FAO), and
(6) \emph{XKOS} data sets (\textcolor{red}{XX}).

The correctness of $\mathcal{C}_{C}$ and $\mathcal{C}_{T}$ constraints, i.e., the gold standard has been designed by a SBE domain expert.
As proof-of-concept, we use the \emph{RDF Validator} to offer examples for both valid and invalid RDF data regarding particular constraints.
